Characterising cheque dishonour cases in India: Causes for delays and policy implications

### 2022-04-07 DOC/paper_chequeDishonour_combined.tex
Discussed edits. Done.

One reason for the large burden on courts is believed to be the large share of \gls{ni} cases.\footnote{In particular, \S~138 of the Negotiable Instruments Act (Dishonour of a cheque for insufficiency, etc., of funds in the account). Act 26 of 1881} {\color{red} Karan: Define NI Act case in 2 lines here.} As per the \gls{lci}, they represent 6.5\% and 7.8\% of all institutions and pendency in Indian courts, respectively \autocite{lci2014_arrears}. As per one order of the Supreme Court of India, they reflect more than 15\% of all criminal cases in the District Courts \autocite{sc2020_makwanavstate}. As per another order, they constitute over 30\% of the total pendency in courts.\footcite[Similarly, a study published by the Department of Justice briefly touches on the burden of such cases on the judiciary and posits that they constitute 34\% of pending criminal cases in Maharashtra.][]{sc2020_138, mahadik2018_maharashtra}

@@

{\color{red}Karan: Add some text here}

>>>>

One reason for the large burden on courts is believed to be the large share of \gls{ni} cases.\footnote{In particular, \S~138 of the Negotiable Instruments Act (Dishonour of a cheque for insufficiency, etc., of funds in the account). Act 26 of 1881} While the NI Act contains several provisions that may lead to a dispute, \S~138, in particular, provides that if a cheque (drawn by a person for paying off a debt or a liability) is returned unpaid due to insufficiency of funds or credit, the payer may be imprisoned for up to two years or with a fine up to twice the amount of the cheque, or both. As per the \gls{lci}, they represent 6.5\% and 7.8\% of all institutions and pendency in Indian courts, respectively \autocite{lci2014_arrears}. As per one order of the Supreme Court of India, they reflect more than 15\% of all criminal cases in the District Courts \autocite{sc2020_makwanavstate}. As per another order, they constitute over 30\% of the total pendency in courts.\footcite[Similarly, a study published by the Department of Justice briefly touches on the burden of such cases on the judiciary and posits that they constitute 34\% of pending criminal cases in Maharashtra.][]{sc2020_138, mahadik2018_maharashtra}

@@

The Supreme Court and the Amici Curiae, while addressing jurisdictional issues, have acknowledged that cases can commence in territorial jurisdiction where the accused does not ordinarily reside. This can lead to delays since Magistrates may not be able to proceed against the accused. They have thus recommended that an inquiry be conducted on receipt of complaints under \S~138 of the Act to arrive at sufficient grounds to proceed against the accused when such accused resides beyond the court's territorial jurisdiction and suitable action be taken as per the CrPC. If courts can address jurisdictional issues, they can significantly reduce the duration of cheque dishonour cases.
